Demystifying the Computer Science \\ PhD Admission in the US\\{\large A Handbook for International \emph{and Domestic} Students

### 2024-03-24 demystify.tex
updates

Short answer: this is highly dependent on the adcom member.  For me, it is likely about 10 minutes on average (less than 5 mins for definite rejection; 15+ mins for an acceptance). Note that my earlier years, I spent more time, but now I have more experience and can quickly tell if an application is good or bad.

@@

I first skim the summary for school, GPA, test scores to looking for red flags like low GPA or test scores that do not meet the minimum requirements. Then I move to transcripts to look for low/failing grades on CS and related courses.  For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ". I do not look at test results as the main scores are already given in the summary page.

@@

Next is the SOP, which I read carefully (of course if it is bad then it would a quick read). My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs. X,Y,Z".  If the SOP mentioned about publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.  I will not go online search for the papers (unless they upload them as sample papers).  I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well". If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has a project on Github with X stars''.

@@

>>>>

Around 15 minutes on average (less than 10 mins for definite rejection; 20+ mins for an acceptance). Note I spent more time in my earlier years, but now I have more experience and can quickly tell if an application is good or bad.

@@

I first skim the summary for school, GPA, test scores to looking for red flags like low GPA or test scores that do not meet the minimum requirements. Then I move to transcripts to look for low/failing grades on CS and related courses.  For these I make comments such as ``- unknown international school with good GPA, too many low grades on CS and related courses" or ``+ very good GPA from top school XYZ". I do not look at test details as the summary already gives the main scores.

@@

Next is the SOP, which I read carefully (but if it is bad then it would a quick read). My notes for SOP include things such as ``- SOP mostly on course work, no clear motivation, generic" , or ``+ very enjoyable SOP talking about XYZ, customized for GMU, mentioned profs. X,Y,Z".  If the SOP mentioned about publications then I would have ``+ has X papers at top venues" or ``+/- has X papers but at unknown places''.  I will not go online search for the papers (unless they upload them as sample papers).  I also skim over sampling papers, essentially looking over the abstract to see if reads well, and make notes such as ``+ have a paper under submission to unknown venue, but reads well". If the student mentioned about a project, I might also look at it (e.g., if it is on Github) and make notes such as ``+ has project on Github with X stars''.

@@

After I submit my evaluation, the adcom chair will look all evaluations for that application and make a decision (almost always following our suggestion).  If there are large discrepancies such as one vote reject but another full funding acceptance (such cases are rare, but do happen), the chair might ask us to coordinate and re-evaluate the application if necessary.
